---
title: "Homework 5"
format: html
editor: visual
---

## Load required libraries and data

```{r}
library(tidyverse)
library(brms)
```

```{r}
hs_dep <- read_csv2("data/hate_depres.csv")
hs_dep
```


The dataset contains data from N = 3827 gays and lesbians. 
- hs - measures how often participants encountered hate speech against gays or lesbians (1 = very rare, 7 = very often)
- depres - measures intensity of depression symptoms (using PHQ-8), 1 = low intensity, 5 = high intensity 
- famAcc - indicates number of family members that accept sexual orientation of participant (from 0 - no one - to 4 - both parents and siblings: brothers and sisters)

1. Fit a model that test effects of hs and famAcc on depres. 

a) Start with standardizing variables

```{r}

```

b) Fit an unconditional model (remember about setting priors)

Because the outcome - depression - is highly skewed you may want to set family to `skew_normal`, which will result in a more appropriate posterior predictive distribution. However, keep in mind that such model may take a little bit longer to fit (up to several minutes).

```{r}

```

b) Fit a conditional model

If you decided to use a `skew_normal` family then you should use the same here.

```{r}

```

c) Decide which model has better predictive accuracy using LOOIC

Usually, with large datasets computing LOOIC may take a while (up to several minutes)

```{r}

```

```{r}

```

```{r}

```

d) Plot the results for each unique value of family acceptance score.

Create a function that will find z-transformed values for scores from 0 to 4, given user defined mean and SD.

```{r}
my_scale2 <- function(x, mean, sd){
  
}
```


Use `conditional_effects` to draw the interaction of z_hs and z_famAcc (z_hs moderated by z_famAcc). 
Use 'int_conditions' to specify on which values of the moderator the z_hs effects should be conditioned.
Use `my_scale2` function (you have to find mean and SD of famAcc and put the values instead of placeholders).


```{r}
conditional_effects(fit_c, 
                    effects = , 
                    int_conditions = list(z_famAcc = my_scale2(0:4, <mean>, <sd>)))
```

e) Use `hypothesis` to find values of effects 'z_hs' for each unique value of family acceptance score

```{r}

```

f) Interpret the results

