---
title: "Homework 4"
format: html
editor: visual
---

## Load the the libraries

```{r}
library(brms)
library(bayestestR)
```

## Load the data

Set the working directory to source file location and load the data with this command.

```{r}
vacc <- read.csv2("../notebooks/data_files/exp_vaccines.csv")
head(vacc)
```

Your task is to find a model that would be the most accurate in predicting vaccination intentions.

-   Use only demographic variables (gender, age, place). Try every combination of these variables (there are 8 combinations in total).
-   Recall that to include more than on predictor in a regression you should separate them with `+` sign, e.g., `ov ~ pred1 + pred2 + pred3`
-   Use a simple approach to regression (with a Normal likelihood, a Normal prior on regression coefficients, and treatment/dummy coding of categorical variables).
-   Remember that you should standardize continuous variables.
-   Choose a model that has the best out-of-sample predictive accuracy.
-   Describe the best model results, and conduct tests on it using a SEXIT framework.

**If you are wondering how to obtain the eight model: Remember that a model without any predictors should also be tested. Such a model is called a Null Model and it contains only the intercept: `vaccination ~ 1`**
